---
title: "Power Calculation in SW-CRT"
author: "Xinlei Chen"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    highlight: pygments
    
---

$$
$$


### Introduction

This Shiny APP was developed under the paper *Sample Size and Power Determination in Stepped Wedge Cluster Randomized Trial Designs with Treatment Transition Periods*. The App allows users to do power calculation under SW-CRT with treatment transition periods. Both the **true model** (model for generating data) and the **fitted model** (model for fitting the generated data) are GLMM model based on the Hussey and Hughes model [1]. Users are allowed to provide their input of each parameter as vector, and the APP will automatically generate the power (as well as model fitting results) using **all combinations of the user inputs**. The results in *.csv* format can be downloaded. 


The APP can be accessed by:

* **(Recommended)** using the following command in R. It allows users to run the APP locally with few limitations.

```{r, eval=F}
library(shiny)
shiny::runGitHub("powerSW-CRT", "antibaseR", ref="main")
```

* clicking the **"APP"** tab in the top menu bar. This cloud version is **NOT** recommended if huge amount of calculation is needed.


$$
$$

### Input Parameter


<style type="text/css">
.tg  {border-collapse:collapse;border-color:#ccc;border-spacing:0;}
.tg td{background-color:#fff;border-color:#ccc;border-style:solid;border-width:1px;color:#333;
  font-family:Arial, sans-serif;font-size:14px;overflow:hidden;padding:10px 5px;word-break:normal;}
.tg th{background-color:#f0f0f0;border-color:#ccc;border-style:solid;border-width:1px;color:#333;
  font-family:Arial, sans-serif;font-size:14px;font-weight:normal;overflow:hidden;padding:10px 5px;word-break:normal;}
.tg .tg-c3ow{border-color:inherit;text-align:center;vertical-align:top}
.tg .tg-7btt{border-color:inherit;font-weight:bold;text-align:center;vertical-align:top}
.tg .tg-0pky{border-color:inherit;text-align:left;vertical-align:top}
</style>
<table class="tg">
<thead>
  <tr>
    <th class="tg-7btt">Parameter</th>
    <th class="tg-7btt">Description</th>
    <th class="tg-7btt">Input</th>
  </tr>
</thead>
<tbody>
  <tr>
    <td class="tg-c3ow">seed</td>
    <td class="tg-0pky">the seed for R's random number generator which helps reproduce the simulation and results</td>
    <td class="tg-c3ow">integer</td>
  </tr>
  <tr>
    <td class="tg-c3ow">cell size</td>
    <td class="tg-0pky">the number of subject per cluster, per group and per time point</td>
    <td class="tg-c3ow">vector</td>
  </tr>
  <tr>
    <td class="tg-c3ow">cluster size</td>
    <td class="tg-0pky">the number of randomization clusters at each time</td>
    <td class="tg-c3ow">vector</td>
  </tr>
  <tr>
    <td class="tg-c3ow">group size</td>
    <td class="tg-0pky">the number of groups within each randomization cluster</td>
    <td class="tg-c3ow">vector</td>
  </tr>
  <tr>
    <td class="tg-c3ow">transition length</td>
    <td class="tg-0pky">the length of the treatment transition period</td>
    <td class="tg-c3ow">vector with value(s) &gt; 0</td>
  </tr>
  <tr>
    <td class="tg-c3ow">beta0 ($\beta_0$)</td>
    <td class="tg-0pky">intercept of the model</td>
    <td class="tg-c3ow">vector</td>
  </tr>
  <tr>
    <td class="tg-c3ow">beta4 ($\beta_4$)</td>
    <td class="tg-0pky">treatment effect of treatment 4</td>
    <td class="tg-c3ow">vector</td>
  </tr>
  <tr>
    <td class="tg-c3ow">rho34 ($\rho_{3:4}$)</td>
    <td class="tg-0pky">treatment effect ration between treatment 3 and 4</td>
    <td class="tg-c3ow">vector</td>
  </tr>
  <tr>
    <td class="tg-c3ow">rho13 ($\rho_{1:3}$)</td>
    <td class="tg-0pky">treatment effect ration between treatment 1 and 3</td>
    <td class="tg-c3ow">vector</td>
  </tr>
  <tr>
    <td class="tg-c3ow">rho24 ($\rho_{2:4}$)</td>
    <td class="tg-0pky">treatment effect ration between treatment 2 and 4</td>
    <td class="tg-c3ow">vector</td>
  </tr>
  <tr>
    <td class="tg-c3ow">time effect size <br>($\tau$ or $\tau_{j}$)</td>
    <td class="tg-0pky">time effect;<br>the number of total time points = cluster size + transition length + 1;<br>if using a non-linear secular trend, please provide the time effect at each time point</td>
    <td class="tg-c3ow">a single 0 indicating no secular trend; <br>a single nonzero value indicating a linear secular trend; <br>a vector with length &gt; 1 indicating a non-linear secular trend</td>
  </tr>
  <tr>
    <td class="tg-c3ow">S.D. of random cluster effect<br>($\sigma^2_{1}$)</td>
    <td class="tg-0pky">the standard deviation of the random cluster effect</td>
    <td class="tg-c3ow">vector with nonnegtive values</td>
  </tr>
  <tr>
    <td class="tg-c3ow"><span style="font-weight:400;font-style:normal">S.D.</span> of random treatment effect<br>($\sigma^2_{2}$)</td>
    <td class="tg-0pky">the standard deviation of the random treatment effect</td>
    <td class="tg-c3ow">vector with nonnegtive values</td>
  </tr>
  <tr>
    <td class="tg-c3ow">iteration</td>
    <td class="tg-0pky">the number of iteration of each simulation</td>
    <td class="tg-c3ow">a single positive integer</td>
  </tr>
  <tr>
    <td class="tg-c3ow">val1 and val2</td>
    <td class="tg-0pky">the APP will automatically conduct 2 test<br>test1: $a \times val1- b \times val2 = 0$<br>test2: $val1 = val2 = 0$</td>
    <td class="tg-c3ow">choose the treatment effect of interests</td>
  </tr>
  <tr>
    <td class="tg-c3ow">linear combination</td>
    <td class="tg-0pky">the linear combination of test 1: $a \times val1- b \times val2 = 0$</td>
    <td class="tg-c3ow">a vector of length 2 with values of a and b</td>
  </tr>
</tbody>
</table>

$$
$$



### Fitted Model

**Model 1: Fixed treatment effect with no secular trend**

$logit(\mu_{ijk}) = \beta_0 + \beta_1*X_{trt_{1_{ij}}} + \beta_2*X_{trt_{2_{ij}}} + \beta_3 * X_{trt_{3_{ij}}} + \beta_4 * X_{trt_{4_{ij}}} + \zeta_{cluster_i}$

**Model 2: Fixed treatment effect with linear secular trend**

$logit(\mu_{ijk}) = \beta_0 + \beta_1*X_{trt_{1_{ij}}} + \beta_2*X_{trt_{2_{ij}}} + \beta_3 * X_{trt_{3_{ij}}} + \beta_4 * X_{trt_{4_{ij}}} + \tau * X_{time_j} + \zeta_{cluster_i}$

**Model 3: Fixed treatment effect with non-linear secular trend**

$logit(\mu_{ijk}) = \beta_0 + \beta_1*X_{trt_{1_{ij}}} + \beta_2*X_{trt_{2_{ij}}} + \beta_3 * X_{trt_{3_{ij}}} + \beta_4 * X_{trt_{4_{ij}}} + \tau_j + \zeta_{cluster_i}$

**Model 4: Random treatment effect with no secular trend**

$logit(\mu_{ijk}) = \beta_0 + \beta_1*X_{trt_{1_{ij}}} + \beta_2*X_{trt_{2_{ij}}} + \beta_3 * X_{trt_{3_{ij}}} + \beta_4 * X_{trt_{4_{ij}}} + \zeta_{cluster_i} + \zeta_{trt_i}$

**Model 5: Random treatment effect with linear secular trend**

$logit(\mu_{ijk}) = \beta_0 + \beta_1*X_{trt_{1_{ij}}} + \beta_2*X_{trt_{2_{ij}}} + \beta_3 * X_{trt_{3_{ij}}} + \beta_4 * X_{trt_{4_{ij}}} + \tau * X_{time_j} + \zeta_{cluster_i} + \zeta_{trt_i}$

**Model 6: Random treatment effect with non-linear secular trend**

$logit(\mu_{ijk}) = \beta_0 + \beta_1*X_{trt_{1_{ij}}} + \beta_2*X_{trt_{2_{ij}}} + \beta_3 * X_{trt_{3_{ij}}} + \beta_4 * X_{trt_{4_{ij}}} + \tau_j + \zeta_{cluster_i} + \zeta_{trt_i}$

$$
$$


### Statistical Test

The program will test the significance of each treatment effect individually. Besides, two hypothesis tests are conducted.

**Test 1:** testing the equivalence of two treatment effects

$H_0: \beta = \beta^{'}$

**Test 2:** testing if two treatment effects equal to zero simutaneously

$H_0: \beta = \beta^{'}=0$


$$
$$

### Output

The output consists of two parts:

* Power
  + power is generated based on the simulation using all combinations of the user input
* Fitted model
  + estimated coefficient, 95% CI and p-value of each treatment
  + estimated coefficient, 95% CI and p-value of time when using linear secular trend in the fitted model
  + 95% CI and p-value of the two hypothesis tests


$$
$$


### Package

The required R packages are:

```{r, eval=F}
library(shiny)
library(tidyverse)
library(plotly)
library(data.table)
library(MASS)
library(shinythemes)
```


$$
$$

### Reference

[1] Hussey, Michael A., and James P. Hughes. "Design and analysis of stepped wedge cluster randomized trials." Contemporary clinical trials 28.2 (2007): 182-191.

$$
$$


